Parallelization of Push-based System for Molecular Simulation Data Analysis with GPU

### 2016-01-24 paper/paperSource.tex
paper changes...

\author{Iliiazbek~Akhmedov, Yi-Cheng~Tu~$\dagger$% <-this % stops a space

%\thanks{* These authors contributed equally to this work.}
\thanks{$\dagger$ Author to whom all correspondence should be sent.}
\thanks{Iliiazbek Akhmedov and Yi-Cheng Tu are with the Department of Computer Science and Engineering, University of South Florida, 4202 E. Fowler Ave., ENB 118, Tampa, FL 33620, U.S.A. Emails:
akhmedovi@mail.usf.edu, ytu@cse.usf.edu}
\thanks{Joseph Fogarty and Sagar Pandit are with the Department of Physics,
University of South Florida, 4202 E. Fowler Ave., ISA 2019, Tampa,
FL 33620, U.S.A. Emails: jcfogart@mail.usf.edu, pandit@usf.edu} }

@@

Scientist gives the properties to simulation software (for example, Gromacs), runs the simulations, and finally get the output file. The output file must then be analyzed to produce certain results which may help him come up with certain consensus on original theoretical model that resulted in the molecular simulation system\cite{Frenkel:api01}. Gromacs is simulation software tool that helps scientist to run the actual simulation. It is a molecular dynamics package primarily designed for biomolecular systems such as proteins and lipids. \cite{Gromacs-online}. Besides the fact that it helps to generate the output files for the simulations, apparently it also helps to analyze the data itself, but the original problem is that it is not as optimized as it can be in order to analyze the data. Gromacs follows approach of pull-pased design, which means that for any given query (e.g. total mass or total charge, which are very similar type of 1-body queries without sophisticated selection) it will pull data separately and generate addition overhead wasting disk read I/O in order to come up with the result.

%%%%%%%%%%%%%%%%%%%%%%%%%%%
In order to achieve objective discovery and to explain the workings
of the scientific systems, scientist must analyze the data produced
by the MS. These analysis oftentimes comprise of computation of very
complex quantities that show statistical properties of the data.
Such queries are of great importance to scientist because they are
the basic assembly blocks for a series of critical quantities needed
to outline the scientific systems \cite{Frenkel:api01}. The methods
by which the data is being accessed as well as these queries are
executed can either increase or decrease the efficiency of an data
analysis system. The known data analysis systems (e.g., Gromacs
\cite{Gromacs-online}) uses pull based type design in which the data
is being fed to the queries only by demand. When a query needs the
data, it requests it, the data is pulled through the system and
delivered to the query. This type of design, we believe, introduces
two types of problematic issues: 1. overhead in I/O traffic by not
allowing for multiple queries to be executed on the same stream of
data at the same time, and 2. cpu/data latency incurred when a
request for the data is sent to the system (the time it takes from a
request to the time the query can actually use the data).

We believe that by using a push-based type system we can remedy such
issues. In the push-based design the queries do not request the
data. Rather, the data is pushed onto the queries automatically and
is being processed by the active queries. This alleviates the I/O
traffic overhead as well as the cpu/data latency incurred for each
query that would have been introduced by a pull-based design. Such a
system design is especially suitable for many scientific
applications which share the following features. First, the
scientific analysis often involves executing a number of analytical
queries commonly processed on a large portion, if not all of the
generated data. Second, many of these fields use the same popular,
and small number of data analytics primitives as the basic blocks on
which they build their discovery. And three, the scientific data,
once stored is never modified. New data only appends to the existing
dataset, making the data perfect contender for streaming.

%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this paper, we describe the idea, design and implementation of
such a push-based system. We would like to point out that the work
presented in this paper is an application of such push-based
strategy that has been previously considered in the database
community in the works presented in~\cite{DataPath,Volcano,Qpipe},
as well as the idea of sharing same tuple of data or data scans
across many queries previously introduced
in~\cite{Candea,PredictablePerformance,CooperativeScans}.

To summarize, data-intensive applications often require compelling
amount of storage space and intensive processing capability, but
also need fast data access and high throughput data analysis.
Therefore, the need of a system that will be optimized to access
data fast, with high throughput, as well as efficiently execute the
analytical queries with a possibility of running multiple queries on
the same data stream is of a great importance to the scientific
community.

@@

In order to achieve objective discovery and to explain the workings
of the scientific systems, scientist must analyze the data produced
by the MS. These analysis oftentimes comprise of computation of very
complex quantities that show statistical properties of the data.
Such queries are of great importance to scientist because they are
the basic assembly blocks for a series of critical quantities needed
to outline the scientific systems \cite{Frenkel:api01}. The methods
by which the data is being accessed as well as these queries are
executed can either increase or decrease the efficiency of an data
analysis system. The known data analysis systems (e.g., Gromacs
\cite{Gromacs-online}) uses pull based type design in which the data
is being fed to the queries only by demand. When a query needs the
data, it requests it, the data is pulled through the system and
delivered to the query. This type of design, we believe, introduces
two types of problematic issues: 1. overhead in I/O traffic by not
allowing for multiple queries to be executed on the same stream of
data at the same time, and 2. cpu/data latency incurred when a
request for the data is sent to the system (the time it takes from a
request to the time the query can actually use the data).

We believe that by using a push-based type system we can remedy such
issues. In the push-based design the queries do not request the
data. Rather, the data is pushed onto the queries automatically and
is being processed by the active queries. This alleviates the I/O
traffic overhead as well as the cpu/data latency incurred for each
query that would have been introduced by a pull-based design. Such a
system design is especially suitable for many scientific
applications which share the following features. First, the
scientific analysis often involves executing a number of analytical
queries commonly processed on a large portion, if not all of the
generated data. Second, many of these fields use the same popular,
and small number of data analytics primitives as the basic blocks on
which they build their discovery. And three, the scientific data,
once stored is never modified. New data only appends to the existing
dataset, making the data perfect contender for streaming.

In this paper, we describe the idea, design and implementation of
such a push-based system. We would like to point out that the work
presented in this paper is an application of such push-based
strategy that has been previously considered in the database
community in the works presented in~\cite{DataPath,Volcano,Qpipe},
as well as the idea of sharing same tuple of data or data scans
across many queries previously introduced
in~\cite{Candea,PredictablePerformance,CooperativeScans}.

To summarize, data-intensive applications often require compelling
amount of storage space and intensive processing capability, but
also need fast data access and high throughput data analysis.
Therefore, the need of a system that will be optimized to access
data fast, with high throughput, as well as efficiently execute the
analytical queries with a possibility of running multiple queries on
the same data stream is of a great importance to the scientific
community.

\begin{figure}
 \centerline{ \includegraphics[width=1\columnwidth]{images/snap-white.eps} }
 \caption{A snapshot of simulated collagen fiber structure}
 \label{fg:snap-white}
\end{figure}

>>>>

\author{Iliiazbek~Akhmedov, Yi-Cheng~Tu~$\dagger$, Vladimir Grupcev, Joseph Fogarty, Sagar Pandit % <-this % stops a space

%\thanks{* These authors contributed equally to this work.}
\thanks{$\dagger$ Author to whom all correspondence should be sent.}
\thanks{Iliiazbek Akhmedov, Vladimir Grupcev, and Yi-Cheng Tu are with the Department of Computer Science and Engineering, University of South Florida, 4202 E. Fowler Ave., ENB 118, Tampa, FL 33620, U.S.A. Emails:
akhmedovi@mail.usf.edu, ytu@cse.usf.edu}
\thanks{Joseph Fogarty and Sagar Pandit are with the Department of Physics,
University of South Florida, 4202 E. Fowler Ave., ISA 2019, Tampa,
FL 33620, U.S.A. Emails: jcfogart@mail.usf.edu, pandit@usf.edu} }

@@

Scientist gives the properties to simulation software (for example, Gromacs), runs the simulations, and finally get the output file. The output file must then be analyzed to produce certain results which may help him come up with certain consensus on original theoretical model that resulted in the molecular simulation system\cite{Frenkel:api01}. Gromacs is simulation software tool that helps scientist to run the actual simulation. It is a molecular dynamics package primarily designed for biomolecular systems such as proteins and lipids. \cite{Gromacs-online}. Besides the fact that it helps to generate the output files for the simulations, apparently it also helps to analyze the data itself, but the original problem is that it is not as optimized as it can be in order to analyze the data. Gromacs follows approach of pull-pased design, which means that for any given query (e.g. total mass or total charge, which are very similar type of 1-body queries without sophisticated selection) it will pull data separately and generate addition overhead wasting disk read I/O in order to come up with the result just for a single query. As it has been proposed in the previous paper, in order to remedy such issues, the push-based design does exactly the opposite, where instead of loading the data on demand for each query, the queries are batched into a network, then the entire dataset is loaded chunk by chunk pushing it through the network which has its internal relationships and dependencies amonth the queries. In this case, since scientific simulation data is run once with specific physical properties, it is never modified, thus, on continuation, it will only append, which means that the processing can also be run on appended frames. \cite{mainPaper}
This type of approach has already been revised by other systems \cite {DataPath,Volcano,Qpipe} as of reusing loaded data through queries produced \cite {Candea,PredictablePerformance,CooperativeScans}.

In this paper, we are incorporating parallelization into the processing part of the proposed design by means of CUDA programming language for GPU. Since storage of the simulation data is very expensive, it might come to the point of analyzing the data on the fly (meaning running the simulation and analyzing it at the same time in a streaming manner), which leads to a problem of optimizing the processing part of the design proposed in the previous paper, because time spent on generating data should be tried to conceal the time spent on the processing part by means of overlapping or simply running it in a quick manner.

@@
